---
title: "STA141B Final Assignment"
author: "Sitong Qian"
date: "12/13/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


For this report,my job is to create a interactive time series plot for covid-19 cases in California county. The data used is extracted from new york time, for the first step, I used professor's code directly to extract data. 
```{r,warning=FALSE,echo=FALSE}
library(XML)
library(RCurl)
library(stringr)
library(httr)
library(ggplot2)
library(plotly)
library(dplyr)
library(MASS)
library(tidyverse)
# We manually found the following link via the Web browser's Developer Tools in the Network tab.
# We found the json files

u = "https://static01.nyt.com/newsgraphics/2020/03/16/coronavirus-maps/a3c88fc852744c0bdb9ff9ee092cc5db785705d6/data/timeseries/en/USA.json"
tt = getURLContent(u, verbose = TRUE, followlocation = TRUE)
library(RJSONIO)
us = fromJSON(tt)
# The URL for the JSON changes each day. So today's URL, we want to find it in the current day's page.
# We get the top-level page
ny = htmlParse(getURLContent("https://www.nytimes.com/interactive/2020/us/coronavirus-us-cases.html?action=click&module=Top%20Stories&pgtype=Homepage"))
# Then we find the HTML elements that have an immediate child text node that contains the string USA.json.
js = xpathSApply(ny, "//*[contains(./text(), 'USA.json')]", xmlValue)
#These are <script> elements containing JavaScript, not JSON.
# But we can find the URL with a regular expression in these
u = gsub('.*"(https://[^"]+USA.json)".*', "\\1", js)
u = unique(u)
# There is only 1 URL repeated multiple times.
# So now we have these
tt = getURLContent(u, verbose = TRUE, followlocation = TRUE)
library(RJSONIO)
us = fromJSON(tt)
length(us$data)


```


Now, I have the covid-19 dataset, my next step is to investigate the information presented in the dataset. I found out the information has a total of 3337 columns since I am aware that the data here is far more than the number of states. I used the sapply function to call out the region type and found out it contains, country-level,state-level, county-level, and territory. Then, I was wondering how to find out the county related to California states, being said, I first google it online, and found out geoid is the unique key here, and for the county in California, all of them start with 06 followed by 3 digits. Thus, I extracted all entries with 06 by a regular expression and I stored the index value as CA_county_index.

```{r,echo=FALSE}
region_type = sapply(1:length(us$data),function(i){
  region_type = us$data[[i]][["geoid"]]
  return(region_type)
  }
)

CA_county_index = grep("USA-06[0-9]", region_type)
CA_county_index <- CA_county_index[1:58]
```

Then, with the CA county index, I now convert and processed the information I needed in the final plot.
First, I fixed the date, since the data is stored by date, I convert the range to date value.
Then, I extracted county name and case and death
For pro-rated cases and deaths, I divided them by population. 
Also, I processed the accumulated cases and deaths to present more meaningful information.


```{r}
## date
date=seq(as.Date("2020-03-01"), as.Date("2020-12-13"), by="days")

## county_name

county_nm = sapply(1:length(us$data),function(i){
  county_name = us$data[[i]][["display_name"]]
  return(county_name)
  }
)

CA_cty = county_nm[CA_county_index]


## case and death
case = sapply(1:length(us$data),function(i){
  case = us$data[[i]][["cases"]]
  return(case)
  }
)
death = sapply(1:length(us$data),function(i){
  death = us$data[[i]][["deaths"]]
  return(death)
  }
)

CA_cases = case[CA_county_index]
CA_deathes = death[CA_county_index]

## Pro-rated rate for cases and deathes
population = sapply(1:length(us$data),function(i){
  population = us$data[[i]][["population"]]
  return(population)
  }
)
CA_population = population[CA_county_index]
Pro_rated_cases = mapply(FUN = `/`, CA_cases, CA_population, SIMPLIFY = FALSE)
Pro_rated_deathes = mapply(FUN = `/`, CA_deathes, CA_population, SIMPLIFY = FALSE)

## Accumulate cases for each date 
acc_case = sapply(1:length(CA_cases),function(i){
  case_daily = CA_cases[[i]]
  acc_case <- case_daily %>% accumulate(`+`)
  return(acc_case)
  }
)



## Accumulate deathes for each date 
acc_death = sapply(1:length(CA_deathes),function(i){
  death_daily = CA_deathes[[i]]
  acc_death <- death_daily %>% accumulate(`+`)
  return(acc_death)
  }
)


```


Since the information I got is a list of list, which can't be used for graphing, I now convert them to list.


```{r}
case_final= sapply(1:length(CA_cases),function(i){
  case_fl = CA_cases[[i]]
  return(case_fl)
  }
)

death_final=sapply(1:length(CA_deathes),function(i){
  death_fl = CA_deathes[[i]]
  return(death_fl)
  }
)

pro_case_final=sapply(1:length(Pro_rated_cases),function(i){
  pro_case_fl = Pro_rated_cases[[i]]
  return(pro_case_fl)
  }
)

pro_death_final=sapply(1:length(Pro_rated_deathes),function(i){
  pro_death_fl = Pro_rated_deathes[[i]]
  return(pro_death_fl)
  }
)



```


Here, I convert everything into a dataframe for interactive plot for later used. 

```{r}
### convert everything into a dataframe for interactive plot
dt_case=as.vector(case_final)
dt_death=as.vector(death_final)
dt_p_case=as.vector(pro_case_final)
dt_p_death=as.vector(pro_death_final)
dt_acc_case=as.vector(acc_case)
dt_acc_death=as.vector(acc_death)
dt_date=rep(date, 58)
dt_county=rep(CA_cty, each = 288)
```


```{r,eval=FALSE}
final_covid_data=as.data.frame(as.vector(as.character(dt_date)))
final_covid_data$case <- dt_case
final_covid_data$death <- dt_death
final_covid_data$pro_case <- dt_p_case
final_covid_data$pro_death <- dt_p_death
final_covid_data$acc_case <- dt_acc_case
final_covid_data$acc_death <- dt_acc_death
final_covid_data$county <- dt_county
colnames(final_covid_data)[1] <- "date"
```

Here, I first graphed it in ggplot with everything listed above and group them by county.
Then, I introduce plotly for creating interactive plot here, with highlight feature and sizingpolicy to make sure that the text I added later will not be covered by the plot.

Next, I added the button by using professor's code, while fixed the function with HideLines.js and hideLines() to make the function works. 


```{r,eval=FALSE}
p1 <- final_covid_data %>%
  highlight_key(~county) %>%
  ggplot(aes(x = date, y = case, color = county, group = county,
             text = paste( "death:", death,
                           "\n case:",case,
                           "\n pro-rated case:",pro_case,
                           "\n pro-rated death:",pro_death,
                           "\n cumulative case:",acc_case,
                           "\n cumulative death:",acc_death))) +
    geom_line() +
    geom_point()


ply1 <- ggplotly(p1, tooltip = c("x", "group", "y", "text"),width = 1000,height = 700) %>%
  highlight(on = "plotly_hover", off = "plotly_doubleclick")

        


  
library(htmltools)
    
h = HTML("<h2>California County Cases</h2><p>
This plot shows the COVID cases for each county in California by date.
<button onclick=\"hideLines()\">Click me</button>")
h2 = HTML("<ol>
<li>Click on the button to hide all of the lines</li>
<li>Hover/mouse over a point to see the tooltip and more details about that day's data for that county</li>
<li>Click on a time series to highlight it and grey out the others</li>
<li>Click on a county name in the legend to toggle the corresponding time series's visibility on
plot</li>
</ol>")

sc = tags$script(paste(readLines("HideLines.js"), collapse = "\n"))
sc = tags$script(src = "HideLines.js")
ply2 = prependContent(ply1, sc, h, h2)
saveWidget(ply2, "combined.html", selfcontained = FALSE)



```


Then, I planed to work on the bonus part, which meant by creating a interactive plot for CA map, to do this, I planned to first search on if R has packages included county-level california map, and then combined such graph with the total's data accumulate cases. But because of time, I have tried numerous way of combining the data into the map, but I failed, I first try add the county in the dataframe, to meet up the requirements for built in map functions. But I falied, next I tried join two dataframe with latitude, but I failed again. 

```{r}
library(usmap)
library(ggmap)
library(rgdal)


acc_case_total <- acc_case[290,]
acc_death_total <- acc_death[290,]
dt_acc_case_total <- as.vector(acc_case_total)
dt_acc_death_total <- as.vector(acc_death_total)
dt_county_bonus <- (CA_cty)
bonus_covid_data=as.data.frame(as.vector(acc_case_total))
bonus_covid_data$case_total <- acc_case_total
bonus_covid_data$death_total <- acc_death_total
bonus_covid_data$fips <- dt_county_bonus
ca_base <- plot_usmap('counties',include = c('CA'))
ca_base

```

